---
title: "2DAR selectivity for Stock Synthesis"
author: "Haikun Xu"
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document demonstrates how to use the stand-alone TMB code mentioned in @xu2019 to estimate the two correlation coefficients for selectivity devs (rho_a and rho_t) outside Stock Synthesis. The SS files for this example code can be found [[here]{.underline}](https://github.com/HaikunXu/2DAR4ss/tree/main/SS%20files).

## Tuning SigmaS

The tuning algorithm is based on Equation 19 in @xu2019. Estimated selectivity devs and their standard errors can be found in the report file. They are saved as a [[csv file]{.underline}](https://github.com/HaikunXu/2DAR4ss/blob/main/SS%20files/Selex_devs.csv) in this example but they can also be directly extracted in R using r4ss. SigmaS_new should be used in new the control file as the new SigmaS until it is close to the value specified in the current control file.

```{r}
Selex_devs <- read.csv("D:/OneDrive - IATTC/Git/2DAR4ss/SS files/Selex_devs.csv", header = TRUE)
head(Selex_devs)

# first check the 2D surface of selectivity devs - they appear to be highly correlated in age and time
library(ggplot2)
Selex_devs$Age <- 2:7
Selex_devs$Time <- rep(1948:2011, each = 6)
ggplot(data = Selex_devs) +
  geom_tile(aes(x = Time, y = Age, fill = Value)) +
  scale_fill_distiller(palette = "RdBu", limits = c(-1,1)*max(abs(Selex_devs$Value))) +
  ggtitle("Selectivity devs") +
  theme_bw()

SigmaS_new <- 
  sqrt(sd(Selex_devs$Value) ^ 2 + sum(Selex_devs$Parm_StDev ^ 2) / nrow(Selex_devs)) # Equation 19
b <- 
  1 - sum(Selex_devs$Parm_StDev ^ 2) / nrow(Selex_devs) / SigmaS_new ^ 2 # Equation 20
```

## Estimating autocorrelation coefficients

The two autocorrelation coefficients can be estimated after the tuning is finished. A stand-alone [[TMB code]{.underline}](https://github.com/HaikunXu/2DAR4ss/blob/main/User%20Manual/2DAR_Selex.cpp) is developed to estimate rho_a and rho_t. You need to install the R package *TMB* to compile the code on your machine.

```{r}
library(TMB)
setwd("D:/OneDrive - IATTC/Git/2DAR4ss/User Manual/")

compile("2DAR_Selex.cpp")
dyn.load("2DAR_Selex.dll")

# format Selex_devs as a matrix
Nages = 6 # 6 age bins
Nyears = 64 # 64 time steps

Selex_devs_matrix <- matrix(Selex_devs$Value, nrow = Nages)

Obj <-
  MakeADFun(
    data = list(
      "Nyears" = Nyears,
      "Nages" = Nages,
      "SDev_hat" = Selex_devs_matrix
    ),
    parameters = list(
      "rho_a" = -log(1 / 0.5 - 1),
      "rho_t" = -log(1 / 0.5 - 1),
      "ln_SigmaS" = log(0.96) # SigmaR is fixed at 0.96 based on the tuning above
    ),
    map = list(ln_SigmaS = factor(NA)), # SigmaR is not estimated here
    random = NULL
  )

Upr = rep(Inf, length(Obj$par))
Lwr = rep(-Inf, length(Obj$par))

Opt = nlminb(
  start = Obj$par,
  objective = Obj$fn,
  gradient = Obj$gr,
  upper = Upr,
  lower = Lwr
)

# rho_a and rho_t are estimated to be:
exp(Opt$par) / (exp(Opt$par) + 1)
```
